\documentclass[Report.tex]{subfiles}
\begin{document}
\section{Morphology}
Morphology is the study of the ways in which words are built up of
smaller \textbf{morphemes}. There are two general units of morphemes:
\begin{itemize}
\item \textbf{stems} - the core meaning-bearing units
\item \textbf{affixes} - bits and pieces that adhere to stems to change
their meanings and grammatical functions
\end{itemize}
For example, the word ``unbelievably'' can be broken down into the
stem `believe' and affixes un-, -able and -ly.

Affixes come in many forms:
\begin{itemize}
	\item \textbf{prefix} - precedes the stem (un-)
	\item \textbf{suffix} - follow after the stem (-able)
	\item \textbf{infix} - inserted into the stem
	\item \textbf{interfix} - linking two words together (speed -o- meter)
	\item \textbf{circumfix} - one part before and one part after
\end{itemize}
English has no real infixes or circumfixes, and there are no real infixes
in other European languages either.


\subsection{Combining morphemes}
There are a few ways of combining morphemes together
\begin{itemize}
	\item \textbf{inflection} - combining the stem with a grammatical
		morpheme, resulting in a word of the same class.
		This may sometimes involve agreement.
		(e.g. I walk, he walks)
	\item \textbf{derivation} - creating a word in another class
		(e.g. computer $\rightarrow$ computerise $\rightarrow$ computerisation)
	\item \textbf{compounding} - combining stems (e.g. dog + house $\rightarrow$ doghouse)
	\item \textbf{clitisation} - attaching a clitic to the stem. A clitic is
		a morpheme that acts as a word, but is typically reduces in form
		and is attached to another word (e.g. we've)
\end{itemize}

\subsubsection{Inflection}
Inflection is mostly productive, i.e. applies to (almost) all words in a class.
Examples of this include verbs in -s form, or -ing participle, plurals
or nouns etc. However, there are some irregular forms such as
catch $\rightarrow$ caught. It is important to remember inflection does
not change the class of a word, only the grammar or agreement.
\subsubsection{Derivation}
In derivation, the class of the words are changed.
\begin{itemize}

\end{itemize}

\subsubsection{Compounding}
\subsubsection{Clitisation}

\end{document}
